\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
I feel extremely priveleged to have had the opportunity to work on a problem of personal interest to me with sufficient financial support to put a roof over my head as well as food (and drink) on the table. To that end I would like to first and foremost take this opportunity to thank the Canadian taxpayers for their important investment in scientific research in these tough economic times, which has funded my research and education on multiple occasions. I hope I will be able to do more to give back in the future.

% Joe
I would like to of course express my gratitude to my supervisor, Joseph Sanderson, who took me into his lab as a complete newbie and who provided me with numerous exciting (and sometimes endlessly frustrating) research experiences as well as for not kicking me out after I broke an expensive neutral density filter wheel (expensive on a grad student's budget). I definitely learnt a lot in my time here.

% Benji, Reza?
I would like to thank my fellow graduate student Benji Wales for being a perpetual source of free research ideas without which this thesis would have looked quite different, and Reza Karimi for being a genuinely cool and always helpful dude.

% Shoja, ...?
I sincerely thank Dr. Shoja'eddin Chenouri of the Department of Statistics \& Actuarial Science and the Statistical Consulting and Collaborative Research Unit for his extensive support and guidance in the development of a rigorous Bayesian inference approach to geometry reconstruction. I regret not working harder to see the project through to the end.

% Jim: for helpful discussions and
I would also like to thank Dr. James Martin for his extensive support and mentorship over the past few years, and for the many discussions which I believe have shaped my interests. I also thank Dr. Wing-Ki Liu, giver of the most difficult undergraduate exams, for generously providing the computational resources which made a large part of this thesis possible. I thank Dr. Marcel Nooijen for originally proposing the idea of using a lookup table to reconstruct molecular geometries which started this research project. I also thank Dr. Donna Strickland for being on my thesis committee and for her willingness to share her infinite yet practical laser wisdom.

% I would like to thank 
I have learned a great deal about science and research from very wise collaborators and through numerous side projects over the past few years (unrelated to this thesis) and from whom I have learned a tidy amount.. To that end, I would like to thank Michael MacDonald and Lucia Zuin (Canadian Light Source, Saskatchewan), Haruo Shiromaru and Jun Matsumoto (Tokyo Metropolitan University), Tomonari Wakabayashi (Kindai University, Osaka), Khaled Ibhrahim (Waterloo Institute for Nanotechnology), and Nhat Ly (Department of Mechanical and Mechatronics Engineering).

% Judy + Anja
Judy McDonnell and Anja Drygala have helped tremendously with the administrative side of being a grad student, thank you!


I would like to thank everyone involved in the production of the television series \emph{The Wire} and \emph{Breaking Bad} for providing excellent and thoughtful procrastination material, without which this thesis would have been finished earlier.

Part of this work was made possible by the high-performance computing facilities and the support staff of the Shared Hierarchical Academic Research Computing Network (SHARCNET: \href{www.sharcnet.ca}{www.sharcnet.ca}) and Compute Canada.

I gratefully acknowledge financial support provided by the Natural Sciences and Engineering Research Council of Canada (NSERC) through their Undergraduate Student Research Awards (USRA) Program and their Canada Graduate Scholarships Master's (CGS M) Program.

\endgroup